---
title: "Investments HW 3"
author: "Zachary Fogelson"
date: "August 9, 2015"
output: pdf_document
---

```{r, include=F}
library(plyr)
library(knitr)
library(pander)
```


### Problem 1


**a**
$$Var(A + B + C + D) = w_a^2Var(A) + w_b^2Var(B) + w_c^2Var(C) + w_d^2Var(D) + $$
  $$ 2 * (w_aw_bcov(A,B) + w_aw_ccov(A,C) + w_aw_dcov(A,D) + w_bw_ccov(B,C) + w_bw_dcov(B, D) + w_cw_dcov(C, D))$$
  
$$Var(Mrkt) = .0715$$


$$Cov(A, A+B+C+D) = Var(A) + Cov(A,B) + C(A,C) + Cov(A,D)$$
Similarly for B,C,D. Interestingly, this corresponds to the sum of the elements in any given row of the Variance Covariance matrix!

$$Cov(A, P) = 0.2308$$
$$Cov(B, P) = 0.2660$$
$$Cov(C, P) = 0.3012$$
$$Cov(D, P) = 0.3364$$

**b**
$$\beta = \frac{Cov(P,A)}{\sigma_P^2}$$

$$\beta_A = 3.226$$
$$\beta_B = 3.718$$
$$\beta_C = 4.210$$
$$\beta_D = 4.702$$

**c**
$$r_A - r_f = \beta_A*[E(r_m) - r_f]$$

$$E(r_A - r_f) = 0.226$$
$$E(r_B - r_f) = 0.260$$
$$E(r_C - r_f) = 0.295$$
$$E(r_D - r_f) = 0.329$$

**d**

MVE Weights of the Portfolio:

A: 24%

B: 25%

C: 25%

D: 26%

The MVE weights are exactly the percentage of each firm over the value of the market.

**e**
MVE Weights of the Portfolio:

A: 25%

B: 25%

C: 25%

D: 25%

The weights have changed.



**f**

MVE Weights of the Portfolio:

A: 29%

B: 23%

C: 24%

D: 24%

Var(P) = .0028

### Problem 2

- Format Data

```{r}
openFormat <- function(x, skips){
  y <- read.csv(x, skip=skips, nrows=1068)
  y <- rename(y, c("X"="Date"))
  y <- y[y$Date > 193200,]
  tmp <- y$Date
  y <- y[,-1]/100
  y$Date <- tmp
  y
}

mrkt <- openFormat("F-F_Research_Data_Factors.CSV", 3)
ports <- openFormat("25_Portfolios_5x5.CSV", 19)
dates <- ports$Date
ports <- ports - mrkt$RF
ports$Date <- dates

```

***a***
```{r, eval=T}
regressors <- names(ports)[-length(ports)]
coefs <- data.frame(Index=c(1,2), row.names = c("alpha", "beta"))
for(i in regressors){
  cs <- coefficients(lm(ports[,i] ~ mrkt$Mkt.RF))
  coefs <- cbind(coefs, rbind(round(as.numeric(cs["(Intercept)"]),5), round(as.numeric(cs["mrkt$Mkt.RF"]),5)))
}
coefs <- coefs[,-1]
names(coefs) <- regressors
# pandoc.table(coefs,split.tables=90 )
```
  
 -------------------------------------------------------------------
   &nbsp;     SMALL.LoBM   ME1.BM2   ME1.BM3   ME1.BM4   SMALL.HiBM 
 ----------- ------------ --------- --------- --------- ------------
  **alpha**    -0.00396   -0.00117   0.00199   0.00443    0.00524   
 
  **beta**      1.651       1.444     1.395     1.274      1.402    
 ------------------------------------------------------------------- 
  
 -------------------------------------------------------------
   &nbsp;     ME2.BM1   ME2.BM2   ME2.BM3   ME2.BM4   ME2.BM5 
 ----------- --------- --------- --------- --------- ---------
  **alpha**  -0.00161   0.0013    0.00259   0.00347   0.00375 
 
  **beta**     1.306     1.262     1.23      1.243     1.394  
 ------------------------------------------------------------- 
  
 -------------------------------------------------------------
   &nbsp;     ME3.BM1   ME3.BM2   ME3.BM3   ME3.BM4   ME3.BM5 
 ----------- --------- --------- --------- --------- ---------
  **alpha**  -0.00082   0.00183   0.00219   0.00293   0.00309 
 
  **beta**     1.281     1.132     1.149     1.18      1.375  
 ------------------------------------------------------------- 
  
 -------------------------------------------------------------
   &nbsp;     ME4.BM1   ME4.BM2   ME4.BM3   ME4.BM4   ME4.BM5 
 ----------- --------- --------- --------- --------- ---------
  **alpha**   0.00011   0.00046   0.0018    0.00245   0.0014  
 
  **beta**     1.085     1.095     1.144     1.159     1.413  
 ------------------------------------------------------------- 
  
 ---------------------------------------------------------------
   &nbsp;     BIG.LoBM   ME5.BM2   ME5.BM3   ME5.BM4   BIG.HiBM 
 ----------- ---------- --------- --------- --------- ----------
  **alpha**   -0.00029    2e-05    0.00084   0.00028   0.00147  
 
  **beta**     0.9492     0.943    0.9748     1.089     1.245   
 ---------------------------------------------------------------
 
**b**
```{r, eval=T}
coefs <- rbind(coefs, sapply(ports, mean))
rownames(coefs)  <- c("alpha", "beta", "actual")
coefs <- data.frame(t(coefs))
kable(coefs)
```


**c**
```{r}
lm.capm <- lm(actual ~ beta, coefs)
summary(lm.capm)
```

**d**
```{r}
coefs$predictions <- mean(mrkt$Mkt.RF) * coefs$beta
plot(coefs$predictions, type="l", col="aquamarine3", ylim = c(0,.05))
lines(coefs$actual, col="coral")
legend("topleft", c("Predicted","Actual"), pch=15, col=c("aquamarine3","coral"), bty="n")
```


### Problem 3

**a**

```{r}
coefs <- data.frame(Index=c(1,2,3,4), row.names = c("alpha", "betaMrkt", "betaSMB", "betaHML"))
for(i in regressors){
  cs <- coefficients(lm(ports[,i] ~ mrkt$Mkt.RF + mrkt$SMB + mrkt$HML))
  coefs <- cbind(coefs, rbind(round(as.numeric(cs["(Intercept)"]),5), round(as.numeric(cs["mrkt$Mkt.RF"]),5), round(as.numeric(cs["mrkt$SMB"]),5), round(as.numeric(cs["mrkt$HML"]),5)))
}
coefs <- coefs[,-1]
names(coefs) <- regressors
```

**b**

```{r}
coefs <- rbind(coefs, sapply(ports, mean))
rownames(coefs)  <- c(rownames(coefs)[-length(rownames(coefs))], "actual")
# pandoc.table(coefs,split.tables=90 )
```
 
 
 ----------------------------------------------------------------------
     &nbsp;      SMALL.LoBM   ME1.BM2   ME1.BM3   ME1.BM4   SMALL.HiBM 
 -------------- ------------ --------- --------- --------- ------------
   **alpha**      -0.00723   -0.00373  -0.00112   0.00105    0.00063   
 
  **betaMrkt**     1.251       1.07      1.051     0.918      0.9758   
 
  **betaSMB**      1.513       1.54      1.227     1.227      1.327    
 
  **betaHML**      0.4489     0.2297    0.5043    0.5848      0.917    
 
   **actual**     0.008049   0.009341   0.01215   0.01371    0.01545   
 ---------------------------------------------------------------------- 
  
 ----------------------------------------------------------------
     &nbsp;      ME2.BM1   ME2.BM2   ME2.BM3   ME2.BM4   ME2.BM5 
 -------------- --------- --------- --------- --------- ---------
   **alpha**    -0.00215  -0.00033   0.00038   0.00062  -0.00029 
 
  **betaMrkt**    1.088     1.027    0.9935    0.9797     1.066  
 
  **betaSMB**     1.136    0.9661    0.8235    0.8167    0.8945  
 
  **betaHML**    -0.2344   0.1497    0.3718     0.568    0.8982  
 
   **actual**   0.007898   0.01048   0.01153   0.01251   0.01389 
 ---------------------------------------------------------------- 
  
 ----------------------------------------------------------------
     &nbsp;      ME3.BM1   ME3.BM2   ME3.BM3   ME3.BM4   ME3.BM5 
 -------------- --------- --------- --------- --------- ---------
   **alpha**    -0.00098   0.00109   0.00061   0.00055  -0.00053 
 
  **betaMrkt**    1.134     1.013     1.007       1       1.116  
 
  **betaSMB**    0.8219    0.5067    0.4327    0.4491    0.5997  
 
  **betaHML**    -0.2384   0.04692   0.3229    0.5565    0.8743  
 
   **actual**   0.008504   0.01007   0.01056   0.01152   0.0131  
 ---------------------------------------------------------------- 
  
 ----------------------------------------------------------------
     &nbsp;      ME4.BM1   ME4.BM2   ME4.BM3   ME4.BM4   ME4.BM5 
 -------------- --------- --------- --------- --------- ---------
   **alpha**     0.00091  -0.00012    4e-04    0.00031  -0.00217 
 
  **betaMrkt**    1.064     1.034     1.048     1.027     1.207  
 
  **betaSMB**    0.3312    0.2162    0.2006    0.2261     0.305  
 
  **betaHML**    -0.3546   0.09609   0.3503    0.5639    0.9602  
 
   **actual**   0.008011  0.008429   0.01013   0.01089   0.01168 
 ---------------------------------------------------------------- 
  
 ------------------------------------------------------------------
     &nbsp;      BIG.LoBM   ME5.BM2   ME5.BM3   ME5.BM4   BIG.HiBM 
 -------------- ---------- --------- --------- --------- ----------
   **alpha**     0.00073    0.00017    7e-05   -0.00166   -0.00164 
 
  **betaMrkt**    1.019     0.9854     0.981     1.034     1.134   
 
  **betaSMB**     -0.15     -0.2122   -0.2313   -0.1686   -0.1353  
 
  **betaHML**    -0.2519    0.0313     0.311    0.6385     0.9772  
 
   **actual**    0.006622  0.006886  0.007933  0.008206   0.01053  
 ------------------------------------------------------------------
 
```{r}
coefs <- data.frame(t(coefs))
kable(coefs)
```

**c**
```{r}
lm.ff3f <- lm(actual ~ betaMrkt + betaSMB + betaHML, coefs)
summary(lm.ff3f)
```

**d**
```{r}
coefs$predictions <- mean(mrkt$Mkt.RF) * coefs$betaMrkt +
  mean(mrkt$SMB) * coefs$betaSMB +
  mean(mrkt$HML) * coefs$betaHML

plot(coefs$predictions, type="l", col="aquamarine3")
lines(coefs$actual, col="coral")
legend("topright", c("Predicted","Actual"), pch=15, col=c("aquamarine3","coral"), bty="n")
```

### Problem 4

**a**

```{r}
covMat <- cov(subset(ports, select = -(Date)))
eigens <- eigen(covMat)
barplot(sqrt(eigens$values))
```

```{r}
loadings <- eigens$vectors[c(1:5),]
pc1 <- matrix(loadings[,1], nrow = 5, ncol = 5)
pc2 <- matrix(loadings[,2], nrow = 5, ncol = 5)
pc3 <- matrix(loadings[,3], nrow = 5, ncol = 5)
pc4 <- matrix(loadings[,4], nrow = 5, ncol = 5)
kable(pc1)
kable(pc2)
kable(pc3)
kable(pc4)
```