---
title: "FTAP Homework 3"
author: "Zachary Fogelson"
date: "July 15, 2015"
output: html_document
---

### Problem 1

```{r, include=F}
library(gdata)
library(plyr)
library(knitr)
library(gplots)
```
**a** Build 95% confidence intervals

```{r}
mfunds <- read.xls(xls = "mfunds (1).xls", sheet = 2)
sds <- sapply(mfunds, sd)
sds <- sds / sqrt(length(mfunds$drefus))
means <- sapply(mfunds, mean)
lower <- means - 2 * sds
upper <- means + 2 * sds
all <- rbind(means, upper, lower)
row.names(all) <- c("means","upper", "lower")
kable(head(all))
barplot2(all["means",], las = 2, plot.ci = T, ci.u = upper, ci.l = lower)
```


**b**
As shown in the plot above, much of the confidence intervals for the Drifus and the Fidel funds
overlap making the sample means not very informative.

### Problem 2

Let $x_i$ be a random variable which is equal to the number of households with a phone.

Assume that the true probability of whether an American family has a phone is .39 and our sample size is 200.

Let Y be a binomial random variable:
$$Y = \Sigma_{i = 1}^{200}(x_i)$$

Because the number of observations in our sample is >30 we assume that the CLT applies and results in
Y being well approximated by a normal distrobution:

$$E[Y] = np = 200 * .39 = 78$$
$$Var[Y] = n*p(1-p) = 47.58$$

***a***

Determine P(X < 70):

```{r}
pnorm(70, mean = 78, sd = sqrt(47.58))
```

***b***

Determine P(X > 90) = 1 - P(X < 91):
```{r}
1 - pnorm(91, mean = 78, sd = sqrt(47.58))
```

### Problem 3

Null Hypothesis:
$H_0: p \le .5$

Alternative:
$H_1: p > .5$

Based on the null hypothesis we assume that I am going to lose. To be confident, I take the highest
value at which I would lose, namely .5.

Sample size: 2000, $\hat{p} = .53$

$$Z = \frac{\hat{p} - p}{\sqrt(\frac{p(1-p)}{2000})}$$

$$Z = \frac{.53 - .5}{\sqrt{\frac{.5^2}{2000}}} = 2.68$$

Therefore, the probability that I get a sample average of .53 with a true population of average
of .5 after sampling 2000 people is the right hand tail of the normal cdf after 2.68.

```{r}
1 - pnorm(2.68, 0, 1)
```

Because .0037 is less than 5% I reject the null hypothesis and with 95% confidence, I expect that I will win the election.


### Problem 4
Goal confidence interval should be less than $+/-2\%$ to any side:

$.02 = 2 * \hat{\sigma}$

$\hat{\sigma} = \sqrt{\frac{p(1-p)}{n}}$

In order to be sure that the confidence interval is not more than .02 assume that $p = .5$ because
that is the maximum of the numerator of the $\hat{\sigma}$ term:

$.02 = 2 * \sqrt{\frac{.5^2}{n}}$

$\frac{1}{4 * .01^2} = n = 2500$

### Problem 5

**a**

```{r}
z <- qnorm(.995)
se <- sqrt(.056*(1-.056) / 1200)
cat("Upperboud: ", .056 + z*se ,"\nLowerbound: ", .056 - z*se)
```

**b**
Null Hypothesis:
$H_0: p = .05$

Alternative:
$H_1: p \ne .05$

$$Z = \frac{\hat{p} - p}{\sqrt(\frac{p(1-p)}{1200})}$$

$$Z = \frac{.056 - .05}{\sqrt(\frac{.05(1-.05)}{1200})}$$

```{r}
z <- .006 / sqrt((.05*(1-.05))/1200)
2 * (1 - pnorm(z))
```

Because the probability of getting a sample with a sample default rate of .056 given a true distrobution
of .05 over 1200 observatoins is >30% and therefore greater than a 5% threshold I fail to reject the null
hypothesis that .05 is the true default rate.


**c**

(Assuming that by "part a" this question means "part b")
My p value for part b was ~.34

**d**
```{r}
z <- .006 / sqrt((.05*(1-.05))/1200)
1 - pnorm(z)
```

Because ~.17 is still greater than .05, I still cannot reject the null hypothesis at a 95% confidence interval.

**e**

Null Hypothesis:
$H_0: p_{Aa} - p_{Bb} < 0$

Alternative:
$H_1: p_{Aa} - p_{Bb} \ge 0$

$$Z = \frac{(\hat{p} - p) - 0}{\sqrt(\frac{p(1-p)}{1200})}$$

$$Z = \frac{(.048 - .056) - 0}{\sqrt(\frac{.056(1-.056)}{1200} + \frac{.048(1-.048)}{1200})}$$

```{r}
z <- ((.048 - .056)  - 0) / sqrt((.056*(1-.056))/1200 + (.048*(1-.048))/1200)
pnorm(z)
```

Supposing that the difference of Aa bonds and Bb bonds are less than 0 then at the likelihood of two samples
.48, .56 is ~18% therefore we cannot reject the null hypothesis that the probability of an Aa default is
less than the probability of a Bb default.

### Problem 6

**a**

Null Hypothesis:
$H_0:$ eye level - bottom shelf = 0

Alternative:
$H_0:$ eye level - bottom shelf $\ne 0$

$$t = \frac{121 - 0}{\frac{344}{\sqrt{36}}}$$

```{r}
t <- (121 - 0) / (344/sqrt(36))
2 * (1 - pt(t, 36 - 1))
```

Because the probability of achieving a difference of 121 sales in 36 weeks given a t distrobution with 35 degrees of freedom is <5% we are confident that we can reject the null hypothesis that there is no difference between the eye level and bottom shelf

**b**

As seen above the pvalue is 0.04203608