---
title: "FTAP Homework 2"
author: "Zachary Fogelson"
date: "July 14, 2015"
output: html_document
---

### Problem 1: Compute Variance and Covariance Matrixes
+ Code:
```{r, include = F}
library(gdata)
library(plyr)
library(knitr)
```

```{r}
countries <- read.xls(xls = "country portfolios.xlsx", sheet = 1)
countries <- rename(x = countries, c("Global.Index.Key...Index.Monthly"="key", "Index.Price...Close.Monthly"="price", "Data.Date...Index.Monthly"="date", "Index.Name"="name"))

indexReturns <- data.frame(seq(1:184))

for(i in unique(countries$name)){
  
  considering <- countries[countries["name"] == i,]
  diff <- considering$price[-1] - considering$price[-length(considering$price)]
  basicReturn <- diff / considering$price[-length(considering$price)]
  indexReturns <- cbind(indexReturns, basicReturn)
  indexReturns <- rename(x = indexReturns, c("basicReturn"=i))
}
drops <- c("seq.1.184.")
indexReturns <- indexReturns[, !(names(indexReturns) %in% drops)]
```

+ Variance Covariance Matrix
```{r}
varCov <- var(indexReturns, y=indexReturns)
kable(head(varCov, n = 5))
```

+ Correlation Matrix
```{r}
corMat <- cor(indexReturns, indexReturns)
kable(head(corMat, n=5))
```

**(a)**
The correlation matrix contains more interpretable information regarding the degree of 
co-movement in each pair of series because it is normalized based on the standard deviations

**(b)**
```{r}
labs <- c("DAX", "FTSE", "Madrid", "Australia", "Nikkei", "Bovespa", "CAC", "Bombay", "IBEX", "S&P")
bar <- barplot(as.matrix(corMat), beside = T, col = terrain.colors(length(names(corMat[,1]))), las = 2, names.arg = labs)
legend("bottomleft",labs, cex=0.6, fill=terrain.colors(10))
```


Based on the barplot one can clearly see that the cac40 and the FTSE 100 and the Composite DAX index are highly correlated, which makes sense given that we saw in class that European economic performance is highly
correlated.


### Problem 2

$x_i$ / $x_{i+1}$ | $x_i = 0$ | $x_i = 1$ | $x_{i+1}$ Marginal|
------------------|-----------|-----------|-------------------|
$x_{i+1} = 0$ | .3 | .2 | .5
$x_{i+1} = 1$ | .2 | .3 | .5
$x_{i}$ Marginal | .5 | .5 | 1

**(a)**
$P(x_{i+1}|x_i) = \frac{.3}{.5} = .6 \ne .3 \therefore$ consecutive trades are dependent.

**(b)**
Trades are positively correlated becaues the probabilities along the diagonal of similar
outcomes between $x_i$ and $x_{i+1}$ are high.

**(c)**
Because the marginal probabilities for $x_i$ and $x_{i+1}$ (seen in the table above) are the same they are identically distributed. Their distrobutions are both (0 -> .5, 1 -> .5)

**(d)** .5

**(e)** $P(x_{i+1} = 1|x_i = 1) = \frac{.3}{.5} = .6$

**(f)** $P(x_{i+1} = 0|x_i = 1) = \frac{.2}{.5} = .4$

**(g)** 

Outcomes | $P(x_{i+1}|x_i=1)$ |
---------|--------------------|
$0$ | .4|
$1$ | .6|

**(h)** 

Outcomes | $P(x_{i+1}|x_i=0)$ |
---------|--------------------|
$0$      | .6                 |
$1$      | .4                 |

**(i)** Buys tend to follow buys and sells tend to follow sells.

**(j)** $P(1,0,0,1,1,1,0) = (.5)(.4)(.6)(.4)(.6)(.6)(.4) = .006912$

**(k)** Conditional on the previous trade having been a buy the probability will increase for the
particular series of buys and sells. $P(1,0,0,1,1,1,0) = (.6)^4(.4)^3 = .008294$

### Problem 3

$y_t = y_{t-1} + x_t$

$y_{t-1} = 100$


**(a)**
No, $y$ is not an iid process because it depends on the previous value in the sequence and therefore the values one would expect for $y$ are not consistent overtime.

**(b)**

Outcomes | $P(y_t)$ |
---------|--------------------|
102      | .1                 |
101      | .25                 |
100      | .3                 |
99      | .25                 |
98     | .1                 |

**(c)**
The distrobution of $y_t$ is the same because we do not have any knowledge besides $y_{t-1}$ and the distrobution of $x_t$.

**(d)**

$P(y_t>101) = .1$

$P(y_t<100) = .35$

$P(y_t>101) + P(y_t<100) = .45$

**(e)**

Outcomes | $P(y_t)$ |
---------|--------------------|
104      | .01                 |
103      | .05                 |
102      | .12                 |
101      | .2                 |
100      | .24                 |
99      | .2                 |
98      | .12                 |
97      | .05                 |
96      | .01                 |

**(f)**

$P(y_t>101) = .18$

$P(y_t<100) = .38$

$P(y_t>101) + P(y_t<100) = .56$

### Problem 4

$x_t$ is a randome variable $\therefore$ 
$$E[c_0 + c_1x_t] = c_0 + c_1 E[x_t]$$
$$Var[c_0 + c_1x_t] = c_1^2 Var[x_t]$$

Because of $y_t$ is a linear function applied to $x_t$ it also is a normal random variable
with a mean of $y_{t-1} + .01$ and $\sigma = .04$ 

($N(y_{t-1} + .01, .04^2)$)

### Problem 5

Model for X

  + X appears to be an iid model. Based on the data set, possible states are:
  
Outcomes | $P(X)$ |
---------|--------------------|
4      | .18                 |
3      | .35                 |
2      | .28                 |
1      | .19                 |


Model for Y

  + Y appears to be a random walk model
  + $y_t = y_{t-1} + z$ where $z$ is an iid
  
Outcomes | $P(z)$ |
---------|--------------------|
-.02      | .13                 |
-.01      | .26                 |
0      | .28                 |
.01      | .24                 |
.02      | .08                 |  


### Problem 6

$x$ / $y$ | $x = 0$ | $x = .05$ | $x = .1$ | $y$ Marginal|
--------------------|-----------|----------|-------------------|
$y = 0$     | .1    | .15         | .05   | .3
$y = .05$   | .15   | .2          | .1    | .45
$y = .1$    | .05   | .1          | .1    | .25
$x$ Marginal | .3   | .45         | .25   | 1

**(a)**

$x$ | $P(x)$
----|------
0   | .3
.05 | .45
.1  | .25

$y$ | $P(y)$
----|------
0   | .3
.05 | .45
.1  | .25

**(b)**

$P(X = 0)  \stackrel{?}{\ne} P(X = 0| Y=0)$

$.3 P(X = 0) \stackrel{?}{\ne} .1/.3$

$.3 \ne 1/3$

The variables are dependent.

**(c)**
Yes

**(d)**
$\mu_{x,y} = .3 * 0 + .45 * .05 + .25 * .1 = .0475$

$Var(x) = E[X^2] + E[X]^2 = .45 * (.05)^2 + .25 *(.1)^2 - .0475^2 = .00136875$

**(e)**

$(0-.0475)*(.1*(0-.0475) + .15*(.05 - -.0475) + .05*(.1-.0475)) + (.05-.0475)*(.15*(0-.0475) \\ + .2*(.05 - -.0475) + .1*(.1-.0475) )+ (.1 -.0475)*(.05*(0-.0475) + .1*(.05 - -.0475) + \\ .1*(.1-.0475)) = .0002406$

**(f)**

$cor(X,Y) = \frac{cov(X,Y)}{S_xS_y} = \frac{.0002406}{.00136875} = .1578$

**(g)**

$Y$ | $P(Y|X=.1)$
----|------
0   | .2
.05 | .4
.1  | .4

**(h)**

$$E[Y|X=.1] = .06$$

**(i)**
The condistional expected value makes sense because when X is high so is Y.

### Problem 7

$Y = \frac{1}{m}S_1 + \ldots + \frac{1}{m}S_m$

$Var(Y) =  \frac{1}{m}^2S_1^2 + \ldots + \frac{1}{m}^2S_m^2 + 2 * [(\frac{1}{m}^2) \rho \sigma^2 \ldots (\frac{1}{m}^2) \rho \sigma^2]$

$Var(Y) =  m(\frac{1}{m}^2S_1^2) + 2 * {m \choose 2}[(\frac{1}{m}^2) \rho \sigma^2]$

$Var(Y) =  \frac{1}{m}S_1^2 + m(m-1) ((\frac{1}{m}^2) \rho \sigma^2)$

$Var(Y) =  \frac{1}{m}S_1^2 + \rho \sigma^2 - \frac{1}{m}\rho \sigma^2$

**(a)**
When m is large $Var(Y) = \rho \sigma^2$

**(b)**
$\sigma^2$

**(c)**
0

### Problem 8

S&P500 | 
Gold   | .2


